---
title: "indra.gsea"
output:
  md_document:
    variant: gfm
---

```{r, include = FALSE}
if (requireNamespace("printr", quietly = TRUE)) {
  library(printr)
  options(printr.digits = 2)  # Set number of significant digits for printr
}

knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# indra.gsea

This package provides a simple interface for Gene Set Enrichment Analysis (GSEA)
using the [INDRA Biomedical Discovery Engine](https://discovery.indra.bio/).
The Biomedical Discovery Engine is built on INDRA CoGEx, a graph database
integrating causal relations, ontological relations, properties, and data,
assembled at scale automatically from the scientific literature and structured
sources.

## Installing

The pre-release version of the package can be pulled from GitHub using the
[remotes](https://cran.r-project.org/package=remotes) package:

```{r install, eval = FALSE}
# install.packages("remotes")
remotes::install_github("clemenshug/indra.gsea")
```

## Usage

Query genes can be either [HGNC](https://www.genenames.org/) gene symbols or
IDs. The package contains an example set of gene IDs related to COVID-19
infection.

```{r usage_example}
library(indra.gsea)

data("covid_related_genes")

head(covid_related_genes)

res <- indra_discrete_gsea(covid_related_genes)
```

The result is a list with the following elements:

```{r}
names(res)
```

For example, enriched Reactome pathways can be accessed as follows:

```{r, include = FALSE}
res$reactome$q <- as.character(signif(res$reactome$q, 2))
res$reactome$p <- as.character(signif(res$reactome$p, 2))
```

```{r}
head(res$reactome, n = 10)
```
