---
title: "indra.gsea"
output:
  md_document:
    variant: gfm
---

```{r, include = FALSE}
if (requireNamespace("printr", quietly = TRUE)) {
  library(printr)
  options(printr.digits = 2)  # Set number of significant digits for printr
}

knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# indra.gsea

This package provides a simple R interface for Gene Set Enrichment
Analysis (GSEA) using the [INDRA Biomedical Discovery
Engine](https://discovery.indra.bio/). The Biomedical Discovery Engine
is built on INDRA CoGEx, a graph database integrating causal relations,
ontological relations, properties, and data, assembled at scale
automatically from the scientific literature and structured sources.

## Installing

The pre-release version of the package can be pulled from GitHub using the
[remotes](https://cran.r-project.org/package=remotes) package:

```{r install, eval = FALSE}
# install.packages("remotes")
remotes::install_github("labsyspharm/r-indra-gsea")
```

## Usage

Query genes can be either [HGNC](https://www.genenames.org/) gene symbols or
IDs. The package contains an example set of gene IDs related to COVID-19
infection.

```{r usage_example}
library(indra.gsea)

data("covid_related_genes")

head(covid_related_genes)

res <- indra_discrete_gsea(covid_related_genes)
```

The result is a list with the following elements:

```{r}
names(res)
```

For example, all entities that are predicted to affect the query genes upstream
are stored in the `indra-upstream` element:

```{r, include = FALSE}
res$`indra-upstream`$q <- as.character(signif(res$`indra-upstream`$q, 2))
res$`indra-upstream`$p <- as.character(signif(res$`indra-upstream`$p, 2))
```

```{r}
head(res[["indra-upstream"]], n = 10)
```

These results are acquired by running overrepresentation analysis using Fisher's
exact test and correcting for multiple comparisons using the Benjamini-Hochberg
(FDR) method with α = 0.05 on genes causally upstream by one step from all
entities in the INDRA database. Analysis was performed using all human genes as
the background set.

## Credits

INDRA Biomedical Discovery Engine was developed by the
[Gyori Lab](https://gyorilab.github.io/) at Northeastern University.

INDRA CoGEx (Context Graph Extension) is an automatically assembled biomedical
knowledge graph which integrates causal mechanisms from INDRA with non-causal
contextual relations including properties, ontologies, and data.

INDRA CoGEx is funded by grant HR00112220036 under the DARPA ASKEM / ARPA-H BDF
programs.
